A Wide-Coverage Symbolic Natural Language Inference System

### 2019-07-14 paper2018/paper.tex
fit
table it is
update nodalida

\begin{abstract}
  We present a system for Natural Language Inference which uses a
  dynamic semantics converter from abstract syntax trees to Coq
  types. It combines the fine-grainedness of a dynamic semantics
  system with the powerfulness of a state-of-the-art proof assistant,
  like Coq. We evaluate the system on all sections of the FraCaS test
  suite, excluding section 6. This is the first system that does a
  complete run on the anaphora and ellipsis sections of the FraCaS. It
  has a better overall accuracy than any previous system.
\end{abstract}
\section{Introduction}
Natural Language Inference (NLI) is the task of determining of whether
an NL hypothesis H follows from an NL premise(s) P. NLI has received a
lot of attention in the Computational Semantics literature and has
been approached using a variety of techniques, ranging from logical
approaches
\citep{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017},
all the way to the recent Deep Learning (DL) models for NLI. The
latter approaches, following a general trend in NLP, have been
dominating NLI and a number of impressive results have been
produced \citep{kim:2018,radford:2018,liu:2019}.\footnote{These are
  the three systems with the best results on SNLI in increasing
  order at the time of writing.}  State-of-the-art DL systems achieve an accuracy of around
0.9 when tested on suitable datasets. However, the datasets that are
used are assuming a definition of inference that can be thought to be
`looser' or less precise compared to the definition assumed in
platforms based in logical approaches \citep{bernardy:2019}. For example, consider the following
 example  from the SNLI dataset, predominatly used to test DL
approaches:
%
\begin{lingex}
\item % An SNLI example \\
  \label{ex:snli}
\textbf{P}  A  man selling donuts to a customer during a world
exhibition event held in the city of Angeles. \\
 \textbf{H}  A woman drinks her coffee in a small cafe.\\
\textbf{Label: Contradiction} [SNLI]
% \item % An example from FraCaS\\
% \textbf{P1} A Scandinavian won the Nobel Prize.\\
% \textbf{P2}	Every Swede is  Scandinavian.\\
% \textbf{Q}  Did a Swede win the Nobel prize? \\
% \textbf{H} A Swede won the Nobel prize.\\
% \textbf{Label: Yes} [FraCaS 173]
\end{lingex}
%
In \ref{ex:snli}, a number of non-trivial
assumptions have to be made in order to arrive at a contradiction: a) the two
situations described have to be taken to refer to the same situation
in order to judge that the latter  contradicts the former, b)
the indefinite article in the premise has to be identified with the
indefinite article in the hypothesis. (Additionally considering that a
person cannot be a man selling donuts and a woman drinking coffee at
the same time.) While this can be part of the reasoning humans
perform, it is not the only possibility. More precise, logical
reasoning is also a possibility, and will render the above label as unknown. Furthermore, reasoning can get very fine-grained 
as the Contained Deletion ellipsis example \ref{ex:acd} below shows:

@@

\paragraph{Coq}
Coq is an interactive theorem prover (proof assistant) based on the
calculus of inductive constructions (CiC), i.e.  a lambda calculus
with dependent types. Coq is one of the leading proof assistants. It
has famously been instrumental in formalising a proof of the
four-color theorem \citep{Gonthier:2008}, a proof of the odd order
theorem \citep{Gonthier:2013}. Coq is a very powerful reasoning engine
that makes it fit for the task of NLI, when the latter is formalized
as theorem proving task. It supports
notably dependent typing and subtyping, which are instrumental in
expressing NL semantics.

@@

\paragraph{Anaphora} Thanks to the above system, we can handle many
anaphoric cases, including E-Type and Donkey pronouns. Indeed, even
objects which have no syntactic representation can be added to the
environment. We follow here the general pattern outlined by
\citet{unger:2011}. However, we use a more general scope-extension
mechanism, which allows us to support examples like  the following:
\begin{lingex}
\item
    \label{ex:anaphora-scope}
\textbf{P1}	Every committee has a chairman.\\
\textbf{P2}	He is appointed its members.\\
\textbf{H} 	Every committee has a chairman appointed by members of the committee. \\ \textbf{Label:	YES} [FraCaS 122]
\end{lingex}
In the above example, the pronoun ``he'' is allowed to refer to the
object quantified over by ``every'', whose scope is extended
accordingly.

@@

\paragraph{Phrasal comparatives}
Previous attempts to tackle the section of the FraCaS test suite
devoted to comparatives showed that handling them is not easy. Our
strategy here is to leverage our dynamic semantics, revealing an
anaphoric element of comparatives. Indeed, Consider the hypothesis of
(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
that this sentence is equivalent to the following two separate parts:
``APCOM lost zero or more orders. ITEL won more orders [than some
elliptic quantity].''  A quantity is introduced every time we talk
about some quantity (indexed by a CN, in this case ``orders''), and it
can be referred to by a comparative, later in the discourse. Using
this idea, we can go one level deeper in the interpretation of our
example: ``APCOM lost $\theta$ orders. $\theta$ $\ge$ 0.  ITEL won at least $\theta$+1
orders.''. We see here how the quantities are introduced. They are
added to the environment so that, they can be referred to as elliptic
quantity expressions.
% SC
%Finally, ``more'' is systematically intepreted as ``at least
%<elliptic quantity>+1''. This treatment, which we illustrated here on
%an example, is systematic in our implementation.

\paragraph{Adjectives}
We interpret gradable adjectives using a pair of a
measure $m : objects \to Z$ and a threshold $\tau : Z$, where $Z$ is
treated as an abstract ordered ring by Coq. (This structure has no
dynamic aspect in our model, and thus is entirely handled within Coq.)
For subsective adjectives, $\tau$ will additionally depend on the class
of the object in question. This structure has the benefit that
opposite adjectives can be easily represented (measures are opposites
$\forall{}x. m_1 x = -m_2 x$ and thresholds do not overlap $\tau_1 + \tau_2 >
0$). Formalization aside, this idea is reminiscent of degree-based
approaches to gradable adjectives of
\citet{Cresswell:1976,Kennedy:2007}.  Additionally
adjectival predicates, as present in the FraCaS suite, are interpreted as
linear inequations in $Z$. Solving systems of such inequations is
decidable.
Indeed, the tactic that Coq offers for this purpose can
solve all such problems in the FraCaS suite, automatically.

@@

FraCaS classifies each problem as either entailment (YES),
entailment of the opposite (NO) or no entailment (UNK).  In this work,
we have amended the FraCaS suite to correct a few problems. First,
certain test case are not formed correctly. Those were already
identified by \citet{maccartney:2007} as such (using an ``undef''
labelling), and we removed those. Second, a few test cases occur twice
in the suite, but with two different labellings (one YES and one UNK),
with an annotation that those labellings correspond to different
readings. However, elsewhere in the suite, if a problem has several
readings but only one has entailment, it occurs only once and is
marked as YES. To make the test suite consistent, if one reading
yields entailment we have always considered it as YES. We have also
removed case 199 (which appears to be vacuous). Finally we changed the
labelling of 6 cases
%(SC: Shall we remove this to the supplementary material and just say to have a look at how we amended the suite?)
% the cases listed in Table \ref{tab:overrulings},
which appeared to have been
misclassified (as justified in supplementary material). We note that the majority of the mistaken
classifications occur in sections 3 and 4, which have not been
previously attempted and thus, we propose, have not been properly
scrutinized.
% \begin{table}
%   \centering
%   \small
% \begin{tabularx}{\columnwidth}{ccp{3.8cm}}
%  test case & new class & comment \\ \hline
%     005 & UNK & missing hypothesis: there are italian tenors \\
%    056 &  Yes & Already identified as such by MacCartney \\
%   069 & Unk & Mary could have used someone else's workstation \\
%   119 & Unk & \textit{ibid.} \\
%   181 & Yes & for the same reason as 180 \\
%   226 & Yes &
% \end{tabularx}
%   \caption{Overruled FraCaS cases}
%   \label{tab:overrulings}
% \end{table}

Our system classifies a case as YES if a proof can be constructed from
the premises to the hypothesis, NO if a proof of the negated
hypothesis can be constructed and UNK otherwise. Because we work with
a non-decidable logic, one cannot \emph{in general} conclude
decisively that no proof exists. Thus, we consider here that no proof
exists if it cannot be constructed with reasonable effort. In
particular, we test at the minimum that the automatic proof search
built in Coq does not succeed before classifying a problem as UNK.

@@

\providecommand\forcecenter{\multicolumn{1}{c}}
\providecommand\ncases[1]{{\ensuremath{^{#1}}}}
\begin{table}
  \centering
  \small
\begin{tabularx}{\columnwidth}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
Section      & {\kern -2em} \#cases & Ours     & FC & MINE & Nut  & Langpro  \\ \hline
Quantifiers  & 75          & .96  & .96    & .77  & .53  & .93  \\
             &             & \ncases{74}     &        &      &      &     \ncases{44} \\
Plurals      & 33          & .82      & .76    & .67  & .52  & .73 \\
      &          &       &     &   &   & \ncases{24} \\
Anaphora     & 28          & .86      &   -    & -    & -    &  -       \\
Ellipsis     & 52          & .87      &   -    & -    & -    &  -       \\
Adjectives   & 22          & .95 & .95    & .68  & .32  & .73 \\
             &          &  \ncases{20} &     &   &   &  \ncases{12} \\
Comparatives {\kern -1em}& 31          & .87      & .56    & .48  & .45  &  -       \\
Temporal     & 75          &  -       &   -    &   -  &  -   &  -       \\
Verbs        & 8           & .75      &   -    & -    & -    &  -       \\
Attitudes    & 13          & .92      & .85    & .77  & .46  & .92  \\ 
    &          &       &     &   &   & \ncases {9}  \\ \hline
Total        & 337         & .89      & .83    & .69  & .50  & .85  \\
             &             & \ncases{259}    & \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
  \end{tabularx}
  \caption{Overview of the accuracy of our system compared to others.
    ``Ours" refers to the approach presented in this paper. When a
    system does not handle the nominal number of test cases (shown in
    the second column), the actual number of test cases attempted is
    shown below the accuracy figure, in smaller font.  ``FraCoq''
    refers to the work of \citet{bernardy_type_2017}. ``MINE" refers
    to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG
    system that utilizes the first-order automated theorem prover
    \textit{nutcracker} described by \citet{bos:2008}, and ``Langpro"
    to the system presented by \citet{Abzianidze:2015}. A dash
    indicates that no attempt was made for a given section. }
  \label{tab:results}
\end{table}

Table \ref{tab:results} shows a considerable improvement over earlier
approaches in terms of coverage, with three more sections covered over
previous state of the art. We thus cover 259 out of 337 cases (77\%),
compared to at most 174 cases (52\%) in previous work. Additionally,
our system performs generally the best in terms of accuracy. In
particular, section 6 sees a large improvement in accuracy, which we
attribute to our new analysis of comparatives, enabled by our dynamic
semantics.

\paragraph{error analysis}
Our system fails to classify correctly 28 cases out of 259.  We give
here a summary of the missing features which are responsible for the
biggest number of failures. The one biggest source of error is an
incomplete handling of group readings. (FraCaS
013, 014, 046, 084, 111, 124, 126, 127, 137, 171, 172, 191, 193, 195, 243, 250, 333, 346). These
are cases where a syntactic conjunction of individuals is treated as a
semantic group, or where precise counting of the members of a group is
necessary.  Other problematic cases are where a definite plural does
not have a universal reading (091, 094, 095) --- these require an
analysis at the pragmatic level, which our system lacks. Additionally,
neither measure phrases (242) nor attributive comparatives (244, 245)
are handled.

\section{Conclusions and Future Work}
We presented a system converting GF trees to Coq types using dynamic
semantics. The system outperforms the state of the art in logical
approaches when tested on the FraCaS test suite and is the only system
to date that has performed a run on the ellipsis/anaphora section of
the FraCaS. The system is precise enough to form the start of a very
precise NL reasoner for controlled domains. For the future, we plan to
extend the system to cover the remaining section of the FraCaS
(tense/aspect), and also develop a more applied version to perform
reasoning on controlled NL domains.

>>>>

\begin{abstract}
  We present a system for Natural Language Inference which uses a
  dynamic semantics converter from abstract syntax trees to Coq
  types. It combines the fine-grainedness of a dynamic semantics
  system with the powerfulness of a state-of-the-art proof assistant. We evaluate the system on all sections of the FraCaS test
  suite, excluding section 6. This is the first system that does a
  complete run on the anaphora and ellipsis sections of the FraCaS. It
  has a better overall accuracy than any previous system.
\end{abstract}
\section{Introduction}
Natural Language Inference (NLI) is the task of determining of whether
an NL hypothesis H follows from an NL premise(s) P. NLI has received a
lot of attention in the Computational Semantics literature and has
been approached using a variety of techniques, ranging from logical
approaches
\citep{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017},
all the way to the recent Deep Learning (DL) models for NLI. The
latter approaches, following a general trend in NLP, have been
dominating NLI and a number of impressive results have been
produced \citep{kim:2018,radford:2018,liu:2019}.\footnote{These are
  the three systems with the best results on SNLI in increasing
  order at the time of writing.}  State-of-the-art DL systems achieve an accuracy of around
0.9 when tested on suitable datasets. However, the datasets that are
used are assuming a definition of inference that can be thought to be
`looser' or less precise compared to the definition assumed in
platforms based in logical approaches \citep{bernardy:2019}. For example, consider the following
 example  from the SNLI dataset, predominatly used to test DL
approaches:
%
\begin{lingex}
\item % An SNLI example \\
  \label{ex:snli}
\textbf{P}  A  man selling donuts to a customer during a world
exhibition event held in the city of Angeles. \\
 \textbf{H}  A woman drinks her coffee in a small cafe.\\
\textbf{Label: Contradiction} [SNLI]
% \item % An example from FraCaS\\
% \textbf{P1} A Scandinavian won the Nobel Prize.\\
% \textbf{P2}	Every Swede is  Scandinavian.\\
% \textbf{Q}  Did a Swede win the Nobel prize? \\
% \textbf{H} A Swede won the Nobel prize.\\
% \textbf{Label: Yes} [FraCaS 173]
\end{lingex}
%
In \ref{ex:snli}, a number of non-trivial
assumptions have to be made in order to arrive at a contradiction: a) the two
situations described have to be taken to refer to the same situation
in order to judge that the latter  contradicts the former, b)
the indefinite article in the premise has to be identified with the
indefinite article in the hypothesis. (Additionally considering that a
person cannot be a man selling donuts and a woman drinking coffee at
the same time.) While this can be part of the reasoning humans
perform, it is not the only possibility. More precise, logical
reasoning is also a possibility, and will render the above label as unknown. Furthermore, reasoning can get very fine-grained 
as the Contained Deletion ellipsis example \ref{ex:acd} below shows:

@@

\paragraph{Coq}
Coq is an interactive theorem prover (proof assistant) based on the
calculus of inductive constructions (CiC), i.e.  a lambda calculus
with dependent types.  Coq is a very powerful reasoning engine
that makes it fit for the task of NLI, when the latter is formalized
as theorem proving task. It supports
notably dependent typing and subtyping, which are instrumental in
expressing NL semantics.

@@

\paragraph{Anaphora} Thanks to the above system, we can handle many
anaphoric cases, including E-Type and Donkey anaphora. Indeed, even
objects which have no syntactic representation can be added to the
environment. We follow here the general monadic semantics approach as outlined by
\citet{unger:2011}. However, we use a more general scope-extension
mechanism, which allows us to support examples like  the following:
\begin{lingex}
\item
    \label{ex:anaphora-scope}
\textbf{P1}	Every committee has a chairman.\\
\textbf{P2}	He is appointed its members.\\
\textbf{H} 	Every committee has a chairman appointed by members of the committee. \\ \textbf{Label:	YES} [FraCaS 122]
\end{lingex}
In the above example, the pronoun ``he'' is allowed to refer to the
object quantified over by ``every'', whose scope is extended
accordingly.

@@

\paragraph{Phrasal comparatives}
Previous attempts to tackle the section of the FraCaS test suite
devoted to comparatives showed that handling them is not easy. Our
strategy here is to leverage our dynamic semantics, revealing an
anaphoric element of comparatives. Indeed, consider the hypothesis of
(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
that this sentence is equivalent to the following two separate parts:
``APCOM lost zero or more orders. ITEL won more orders [than some
elliptic quantity].''  A quantity is introduced every time we talk
about some quantity (indexed by a CN, in this case ``orders''), and it
can be referred to by a comparative, later in the discourse. Using
this idea, we can go one level deeper in the interpretation of our
example: ``APCOM lost $\theta$ orders. $\theta$ $\ge$ 0.  ITEL won at least $\theta$+1
orders.''. We see here how the quantities are introduced. They are
added to the environment so that, they can be referred to as elliptic
quantity expressions.\footnote{The degree parameter assumption is not new in the formal semantics literature \cite{Cresswell:1976,Heim:2001,Kennedy:2007,chatzikyriakidis:2017} among many others. The specific details and computational implementation, however, are.}
% SC
%Finally, ``more'' is systematically intepreted as ``at least
%<elliptic quantity>+1''. This treatment, which we illustrated here on
%an example, is systematic in our implementation.

\paragraph{Adjectives}
We interpret gradable adjectives using a pair of a
measure $m : objects \to Z$ and a threshold $\tau : Z$, where $Z$ is
treated as an abstract ordered ring by Coq. (This structure has no
dynamic aspect in our model, and thus is entirely handled within Coq.)
For subsective adjectives, $\tau$ will additionally depend on the class
of the object in question. This structure has the benefit that
opposite adjectives can be easily represented (measures are opposites
$\forall{}x. m_1 x = \lnot m_2 x$ and thresholds do not overlap $\tau_1 + \tau_2 >
0$). Formalization aside, this idea is reminiscent of degree-based
approaches to gradable adjectives of
\citet{Cresswell:1976,Kennedy:2007}.  Additionally
adjectival predicates, as present in the FraCaS suite, are interpreted as
linear inequations in $Z$. Solving systems of such inequations is
decidable.
Indeed, the tactic that Coq offers for this purpose can
solve all such problems in the FraCaS suite, automatically.

@@

FraCaS classifies each problem as either entailment (YES),
entailment of the opposite (NO) or no entailment (UNK).  In this work,
we have amended the FraCaS suite to correct a few problems. First,
certain test case are not formed correctly. Those were already
identified by \citet{maccartney:2007} as such (using an ``undef''
labelling), and we removed those. Second, a few test cases occur twice
in the suite, but with two different labellings (one YES and one UNK),
with an annotation that those labellings correspond to different
readings. However, elsewhere in the suite, if a problem has several
readings but only one has entailment, it occurs only once and is
marked as YES. To make the test suite consistent, if one reading
yields entailment we have always considered it as YES. We have also
removed case 199 (which appears to be vacuous). Finally we changed the
labelling of 6 cases
%(SC: Shall we remove this to the supplementary material and just say to have a look at how we amended the suite?)
% the cases listed in Table \ref{tab:overrulings},
which appeared to have been
misclassified. We note that the majority of the mistaken
classifications occur in sections 3 and 4, which have not been
previously attempted and thus, we propose, have not been properly
scrutinized. In terms of comparison, this only has a minor effect, since our system is the first system to run sections 3 and 4. 
% \begin{table}
%   \centering
%   \small
% \begin{tabularx}{\columnwidth}{ccp{3.8cm}}
%  test case & new class & comment \\ \hline
%     005 & UNK & missing hypothesis: there are italian tenors \\
%    056 &  Yes & Already identified as such by MacCartney \\
%   069 & Unk & Mary could have used someone else's workstation \\
%   119 & Unk & \textit{ibid.} \\
%   181 & Yes & for the same reason as 180 \\
%   226 & Yes &
% \end{tabularx}
%   \caption{Overruled FraCaS cases}
%   \label{tab:overrulings}
% \end{table}

Our system classifies a case as YES if a proof can be constructed from
the premises to the hypothesis, NO if a proof of the negated
hypothesis can be constructed and UNK otherwise. Because we work with
a non-decidable logic, one cannot \emph{in general} conclude
decisively that no proof exists. Thus, we consider here that no proof
exists if it cannot be constructed with reasonable effort. In
particular, we test at the minimum that the automatic proof search
built in Coq does not succeed before classifying a problem as UNK.\footnote{The other way this can be done is by introducing a timeout as \citet{Mineshima:2015} have done.}

@@

\providecommand\forcecenter{\multicolumn{1}{c}}
\providecommand\ncases[1]{{\ensuremath{^{#1}}}}
\begin{table}[hbt]
  \centering
  \small
\begin{tabularx}{\columnwidth}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
Section      & {\kern -2em} \#cases & Ours     & FC & MINE & Nut  & Langpro  \\ \hline
Quantifiers  & 75          & .96  & .96    & .77  & .53  & .93  \\
             &             & \ncases{74}     &        &      &      &     \ncases{44} \\
Plurals      & 33          & .82      & .76    & .67  & .52  & .73 \\
      &          &       &     &   &   & \ncases{24} \\
Anaphora     & 28          & .86      &   -    & -    & -    &  -       \\
Ellipsis     & 52          & .87      &   -    & -    & -    &  -       \\
Adjectives   & 22          & .95 & .95    & .68  & .32  & .73 \\
             &          &  \ncases{20} &     &   &   &  \ncases{12} \\
Comparatives {\kern -1em}& 31          & .87      & .56    & .48  & .45  &  -       \\
Temporal     & 75          &  -       &   -    &   -  &  -   &  -       \\
Verbs        & 8           & .75      &   -    & -    & -    &  -       \\
Attitudes    & 13          & .92      & .85    & .77  & .46  & .92  \\ 
    &          &       &     &   &   & \ncases {9}  \\ \hline
Total        & 337         & .89      & .83    & .69  & .50  & .85  \\
             &             & \ncases{259}    & \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
  \end{tabularx}
  \caption{Accuracy of our system compared to others.
    ``Ours" refers to the approach presented in this paper. When a
    system does not handle the nominal number of test cases (shown in
    the second column), the actual number of test cases attempted is
    shown below the accuracy figure, in smaller font.  ``FraCoq''
    refers to the work of \citet{bernardy_type_2017}. ``MINE" refers
    to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG
    system that utilizes the first-order automated theorem prover
    \textit{nutcracker} \cite{bos:2008}, and ``Langpro"
    to the system presented by \citet{Abzianidze:2015}. A dash
    indicates that no attempt was made for the section. }
  \label{tab:results}
\end{table}

Table \ref{tab:results} shows a considerable improvement over earlier
approaches in terms of coverage, with three more sections covered over
previous approaches. We thus cover 259 out of 337 cases (77\%),
compared to at most 174 cases (52\%) in previous work. Additionally,
our system performs generally the best in terms of accuracy. In
particular, section 6 largely improves in accuracy, which we
attribute to our dynamic semantics analysis of comparatives.

\paragraph{error analysis}
Our system fails to correctly classify  28 cases out of 259.  We give
here a summary of the missing features which are responsible for the failures. The  biggest source of error is 
incomplete handling of group readings. (FraCaS
013, 014, 046, 084, 111, 124, 126, 127, 137, 171, 172, 191, 193, 195, 243, 250, 333, 346). These
are cases where  a syntactic conjunction of individuals is treated as a
semantic group, or where precise counting of the members of a group is
necessary.  Other problematic cases include definite plurals with
no  universal readings (091, 094, 095). Additionally,
neither measure phrases (242) nor attributive comparatives (244, 245)
are handled.

\section{Conclusions and Future Work}
We presented a system converting GF trees to Coq types using dynamic
semantics. The system outperforms the state of the art in logical
approaches when tested on the FraCaS  and is the only system
to date to perform a run on the FraCaS ellipsis/anaphora section. The system is precise enough to form the start of a 
precise NL reasoner for controlled domains. In the future, we plan to
extend the system to cover the remaining section of the FraCaS
(tense/aspect), and also develop a more applied version to perform
reasoning on controlled NL domains.
